Timing and magnitude of climate driven range shifts in transboundary fish stocks challenge their management

### 2021-11-25 Scripts/Initial_Analysis.Rmd
Updated manuscript and analysis to inlcude zooms to latin america and south pacific islands as supplements

lat_ammap <- ggplot(toe_map_data) +
    geom_sf(
      aes(
        fill = reverse #original map
        # fill = value
      ),
      color =  "white",
      size = 0.1
    ) +
    # Inlcude the participation of species
    geom_sf(data = World_map_part,
            # aes(fill = participation+2000), #original map
            aes(fill = participation, 
                color = participation),
            # color =  "white",
            size = 0.1
    ) +
    # grey maps
    # ggplot() +
    geom_sf(data = grey_land, aes(), fill = "grey50", 
            # color =  "white",
            size = 0.1) +
    # geom_sf(data = grey_eez, aes(), fill = "white", color =  "grey50",size = 0.1) +
    geom_sf(data = no_toe_eez, aes(), fill = "lightblue", color =  "white",size = 0.1) +
    geom_sf(data = no_toe_land, aes(), fill = "lightblue", color =  "white",size = 0.1) +
  coord_sf(xlim = c(-9000000,-5000000),ylim = c(1000000, 4000000)) + 
  facet_wrap(~tresh) +
  scale_colour_viridis("",
                         limits = c(2000,2100),
                         breaks = seq(2000,2100,25),
                         # labels = c("","","","",""),
                         direction = 1
    ) +
    scale_fill_viridis("",
                         limits = c(2000,2100),
                         breaks = seq(2000,2100,25),
                         # labels = c("","","","",""),
                         direction = 1
    ) +
    my_ggtheme_m(map_type = "regional",leg_pos = "")

ggsave("lat_am_map.png",
       lat_ammap,
       height = 5,
       width = 6)

@@

  sp_map <-  ggplot(spi_toe_map_data) +
    geom_sf(
      aes(
        fill = reverse #original map
        # fill = value
      ),
      color =  "white",
      size = 0.1
    ) +
    # Inlcude the participation of species
    geom_sf(data = spi_map_part,
            # aes(fill = participation+2000), #original map
            aes(fill = participation, 
                color = participation),
            # color =  "white",
            size = 0.1
    ) +
    # grey maps
    geom_sf(data = spi_no_toe_eez, aes(), fill = "lightblue", color =  "white",size = 0.1) +
    geom_sf(data = spi_no_toe_land, aes(), fill = "lightblue", color =  "white",size = 0.1) +
    coord_sf(xlim = c(9500000,1000000),ylim = c(-3500000, 1900000)) + 
    facet_wrap(~tresh) +
    scale_colour_viridis("",
                         limits = c(2000,2100),
                         breaks = seq(2000,2100,25),
                         # labels = c("","","","",""),
                         direction = 1
    ) +
    scale_fill_viridis("",
                         limits = c(2000,2100),
                         breaks = seq(2000,2100,25),
                         # labels = c("","","","",""),
                         direction = 1
    ) +
    my_ggtheme_m(map_type = "regional",leg_pos = "bottom") +
    theme(legend.key.width = unit(3, "line"))

>>>>

lat_am_map <- ggplot(toe_map_data) +
    geom_sf(
      aes(
        fill = reverse #original map
        # fill = value
      ),
      color =  "white",
      size = 0.1
    ) +
    # Inlcude the participation of species
    geom_sf(data = World_map_part,
            # aes(fill = participation+2000), #original map
            aes(fill = participation, 
                color = participation),
            # color =  "white",
            size = 0.1
    ) +
    # grey maps
    # ggplot() +
    geom_sf(data = grey_land, aes(), fill = "grey50", 
            # color =  "white",
            size = 0.1) +
    # geom_sf(data = grey_eez, aes(), fill = "white", color =  "grey50",size = 0.1) +
    geom_sf(data = no_toe_eez, aes(), fill = "lightblue", color =  "white",size = 0.1) +
    geom_sf(data = no_toe_land, aes(), fill = "lightblue", color =  "white",size = 0.1) +
  coord_sf(xlim = c(-9000000,-5000000),ylim = c(1000000, 4000000)) + 
  facet_wrap(~tresh) +
  scale_colour_viridis("",
                         limits = c(2000,2100),
                         breaks = seq(2000,2100,25),
                         # labels = c("","","","",""),
                         direction = 1
    ) +
    scale_fill_viridis("",
                         limits = c(2000,2100),
                         breaks = seq(2000,2100,25),
                         # labels = c("","","","",""),
                         direction = 1
    ) +
    my_ggtheme_m(map_type = "regional",leg_pos = "")

ggsave("lat_am_map.png",
       lat_am_map,
       width = 8,
       height = 6)

@@

spi_map <-  ggplot(spi_toe_map_data) +
  geom_sf(
    aes(
      fill = reverse #original map
      # fill = value
    ),
    color =  "white",
    size = 0.1
  ) +
  # Inlcude the participation of species
  geom_sf(data = spi_map_part,
          # aes(fill = participation+2000), #original map
          aes(fill = participation, 
              color = participation),
          # color =  "white",
          size = 0.1
  ) +
  # grey maps
  geom_sf(data = spi_no_toe_eez, aes(), fill = "lightblue", color =  "white",size = 0.1) +
  geom_sf(data = spi_no_toe_land, aes(), fill = "lightblue", color =  "white",size = 0.1) +
  coord_sf(xlim = c(9500000,1000000),ylim = c(-3500000, 1900000)) + 
  facet_wrap(~tresh) +
  scale_colour_viridis("",
                       limits = c(2000,2100),
                       breaks = seq(2000,2100,25),
                       # labels = c("","","","",""),
                       direction = 1
  ) +
  scale_fill_viridis("",
                     limits = c(2000,2100),
                     breaks = seq(2000,2100,25),
                     # labels = c("","","","",""),
                     direction = 1
  ) +
  scale_x_continuous(limits = c(-120,180),
                     breaks = c(-120,180)
                     # labels =c("a","b")
                     ) + 
  my_ggtheme_m(map_type = "regional",leg_pos = "bottom") +
  theme(legend.key.width = unit(3, "line"))

ggsave("spi_map.png",
       spi_map,
       width = 8,
       height = 6)
```

### 2021-11-25 Scripts/Manuscript.Rmd
Updated manuscript and analysis to inlcude zooms to latin america and south pacific islands as supplements

The median year in which transboundary stocks experience a range shift varied significantly according to the species’ habitat association (Methods; Kruskal-Wallis, $X^2$ = 203.85, DF = 93, *p* < 0.001; using the 2 s.d.  threshold; Tables \@ref(tab:toeTableS1), \@ref(tab:toeTableS2) and Fig. \@ref(fig:toeFigureS6)) and EEZs' geographic regions (Methods; Kruskal-Wallis, $X^2$ = 242.11, DF = 93, *p* < 0.001; using the 2 s.d.  threshold; Table S3 and Fig. 2A). Overall, most tropical EEZs and stocks will see earlier shifts, with the EEZs of Latin America, the Caribbean and Polynesia experiencing shifts significantly earlier (*p* < 0.05; Tables \@ref(tab:toeTableS2) and \@ref(tab:toeTableS3) than almost any other region (Figs. \@ref(fig:toeFigure2)B, \@ref(fig:toeFigureS3) and \@ref(fig:toeFigureS4)). In contrast, EEZs and stocks located in temperate regions, like northern Europe and eastern Asia, are projected to experience much later range shifts. Stocks in a number of EEZs including Brazil and New Zealand’s EEZs are projected to experience no range shifts beyond natural variability by the end of this century (Fig. \@ref(fig:toeFigure2)A).

@@

```{r toeFigure2, eval = T,  echo = F, fig.width  = 13,fig.pos= "H", fig.height =11, fig.cap =  "Year of change in the shared distribution of 4,119 transboundary stocks. A) Land polygons show the contribution of shifting stocks to a country or territory’s total fishing revenue from transboundary stocks. Exclusive Economic Zone polygons display average year of range shifts within them. Warm colors are indicative of an early shift/high fishing revenue contribution from transboundary stocks while cold colors represent a late shift/low fishing revenue. EEZs with no distributional shift between 2006 and 2100 are represented in pale blue. B) Year of shared distribution shifts by UN sub-regions. Points color coded by the median year for all shifting stocks in the region. Horizontal dashed line represents the year by which countries have committed to reach full implementation of the 2030 Agenda. N = North, S = South, W = West and E = East. Ltn. Ame. and the Car. = Latin America and the Caribbean. Aus and New Z. = Australia and New Zealand."}

>>>>

The median year in which transboundary stocks experience a range shift varied significantly according to the species’ habitat association (Methods; Kruskal-Wallis, $X^2$ = 203.85, DF = 93, *p* < 0.001; using the 2 s.d.  threshold; Tables \@ref(tab:toeTableS1), \@ref(tab:toeTableS2) and Fig. \@ref(fig:toeFigureS6)) and EEZs' geographic regions (Methods; Kruskal-Wallis, $X^2$ = 242.11, DF = 93, *p* < 0.001; using the 2 s.d.  threshold; Table S3 and Fig. 2A). Overall, most tropical EEZs and stocks will see earlier shifts, with the EEZs of Latin America, the Caribbean, Melanesia and Polynesia experiencing shifts significantly earlier (*p* < 0.05; Tables \@ref(tab:toeTableS2) and \@ref(tab:toeTableS3) than almost any other region (Figs. \@ref(fig:toeFigure2)B, \@ref(fig:toeFigureS3) and \@ref(fig:toeFigureS4)). In contrast, EEZs and stocks located in temperate regions, like northern Europe and eastern Asia, are projected to experience much later range shifts. Stocks in a number of EEZs including Brazil and New Zealand’s EEZs are projected to experience no range shifts beyond natural variability by the end of this century (Fig. \@ref(fig:toeFigure2)A).

@@

```{r toeFigure2, eval = T,  echo = F, fig.width  = 13,fig.pos= "H", fig.height =11, fig.cap =  "Year of change in the shared distribution of 4,119 transboundary stocks. A) Land polygons show the contribution of shifting stocks to a country or territory’s total fishing revenue from transboundary stocks. Exclusive Economic Zone polygons display average year of range shifts within them. Warm colors are indicative of an early shift/high fishing revenue contribution from transboundary stocks while cold colors represent a late shift/low fishing revenue. EEZs with no distributional shift between 2006 and 2100 are represented in pale blue. See Fig. S11 for zoom into the Caribbean and South Pacific Islands.  B) Year of shared distribution shifts by UN sub-regions. Points color coded by the median year for all shifting stocks in the region. Horizontal dashed line represents the year by which countries have committed to reach full implementation of the 2030 Agenda. N = North, S = South, W = West and E = East. Ltn. Ame. and the Car. = Latin America and the Caribbean. Aus and New Z. = Australia and New Zealand."}
